---
title: "Trends in harms associated with `r params$hdrug` in `r paste0(params$jur)`, `r params$yr97[1]`
  - `r params$yr97[2]`"
author: "Drug Trends"
output:
  pdf_document: default
  html_document: default
params:
  yr97: NA
  jur: NA
  hdrug: NA
  hospital: hospitalReactive()
  treatment: treatmentReactive()
  jurdata: jurdata()
---

This page contains data on drug harms in Australia from various sources, including the National Hospital Morbidity Database (NHMD), Cause of Death Unit Record File (COD-URF) and The National Alcohol and Other Drug Treatment Services National Minimum Data Set (AODTS NMDS). Principal diagnosis was used to identify drug-induced deaths and hospitalizations. The menu option 'All drugs' is the sum of all drugs within the datasets and includes drugs that are currently unavailable in isolation.

```{r, echo = FALSE, include=FALSE}
library(shiny)
library(plotly)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(haven)
```

## Drug-induced Hospitalisations

This plot depicts the rate of hospitalisations caused by `r params$hdrug` in `r params$jur` from `r params$yr97[1]` - `r params$yr97[2]`. The mean rate for all genders in `r params$yr97[2]` was `r paste0(round(tail(params$hospital$cr, n=1),digits=1))` per 100,000 people.

```{r, echo = FALSE, warning=FALSE}
p <- ggplot(params$hospital) + aes(x = year, y = cr, colour = Sex, linetype = Sex, group = 1, text = paste0(
                      "Year: ", year,
                      "<br>Location: ", jurisdiction,
                      "<br>Drug: ", Drug,
                      "<br>Hospitalisations: ", n,
                      "<br>Crude Rate: ", cr_p
                    )) +
       geom_line() + geom_point(size=0.55) +
       labs(x = "Financial Year", colour = "", linetype = "") +
       scale_color_manual(values = c("red", "blue", "black")) +
       scale_linetype_manual(values = c("dashed", "dotdash", "solid")) +
       theme(panel.grid.major.x = element_blank(),
             panel.grid.major.y = element_line(color = "gray90", linewidth = 0.2),
             panel.grid.minor.y = element_line(color = "gray90", linewidth = 0.2),
             panel.background = element_blank()) +
       scale_x_continuous(breaks = c(2000, 2005, 2010, 2015, 2020),
                          labels=c("2000", '2005', '2010', '2015', '2020')) + 
       scale_y_continuous(limits = c(0, max(params$hospital$cr_uci, 2.5))) +
         labs(x = "Financial Year", y = "Crude Rate per 100,000")
ggplotly(p, tooltip = "text")
```

## Drug-induced Deaths

This plot depicts the rate per 100,000 people for deaths where `r params$hdrug` was the principal cause. These deaths were reported in `r params$jur` and the mean rate was `r paste0(round(tail(params$jurdata$cr, n=1), digits = 1))` per 100,000 people in `r params$yr97[2]`.

```{r, echo = FALSE, warning=FALSE}

gp <- ggplot(params$jurdata, aes(x = Year, y= cr, colour = Sex, linetype = Sex, group = 1, text = paste0(
      "Year: ", Year,
      "<br>Location: ", jurisdiction,
      "<br>Drug: ", Drug,
      "<br>Deaths: ", n,
      "<br>Crude Rate: ", cr_p
    ))) + geom_point(size=0.55) +
    labs(x = "Year", colour = "", linetype = "") + 
      scale_color_manual(values = c("red", "blue", "black")) +
      scale_linetype_manual(values = c("dashed", "dotdash", "solid")) +
      theme_light() + geom_line() + 
      scale_y_continuous(limits = c(0, max(params$jurdata$cr_uci, 2.5))) +
      labs(x = "Year", y = "Crude Rate per 100,000")
    ggplotly(gp, tooltip="text")

```

## Use of illicit drug treatment services
                       
Treatment services in `r params$jur` for `r params$hdrug` were utilized at a rate of `r paste0(round(tail(params$treatment$cr, n=1), digits = 1))` per 100,000 people in `r params$yr97[2]`. The rate when the current reporting period began was `r paste0(round(head(params$treatment$cr, n=1), digits = 1))` in `r params$yr97[1]`.

```{r, echo = FALSE, warning=FALSE}
gp <- ggplot(params$treatment) + aes(x = Year_end, y = cr, colour = Sex, linetype = Sex, group = 1, 
                          text = paste0(
                            "Year ", Year_end,
                            "<br>Crude rate: ", round(cr, digits = 3),
                            "<br>Location: ", Location,
                            "<br>Drug: ", Drug
                          )) + 
      geom_point(size=0.55) +
      labs(x = "Year", colour = "", linetype="") + 
      scale_color_manual(values = c("red", "blue", "black")) +
      scale_linetype_manual(values = c("dashed", "dotdash", "solid")) +
      theme(panel.grid.major.x = element_blank(),
            panel.grid.major.y = element_line(color = "gray90", linewidth = 0.2),
            panel.grid.minor.y = element_line(color = "gray90", linewidth = 0.2),
            panel.background = element_blank())+ geom_line() + 
        scale_y_continuous(limits = c(0, max(params$treatment$cr_uci, 2.5))) +
        labs(x = "Year", y = "Crude Rate per 100,000")
    ggplotly(gp, tooltip="text")

```


